Restrictive Clearance Comparison

### 2024-11-04 data-raw/RestrictiveClearance_comparison.Rmd
Updated object 'cvt' and related documentation. invivPK_loq_units were not ug/mL as intended. Added description of each column of the data. Added description for fup-optimizing clearance aware 1-compartment model in data.R

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

```{r library_loading}
devtools::load_all()
library(httk)
library(tidyverse)
my_pk <- readRDS("data-raw/all_cvt_pk_JOINT.rds")
```

@@

```{r chems_in_common}
common_chems <- cvt %>%
  subset(analyte_dtxsid %in% get_cheminfo(info = "DTXSID", model = "gas_pbtk")) %>%
  pull(analyte_dtxsid) %>%
  unique()

common_chems[!(common_chems %in% get_cheminfo(info = "DTXSID", model = "gas_pbtk"))] %>%
  unique() %>%
  sort()
```

```{r dawson}
# try dawson
httk::load_dawson2021()

common_chems_dawson <- cvt %>%
  subset(analyte_dtxsid %in% get_cheminfo(info = "DTXSID", model = "gas_pbtk")) %>%
  pull(analyte_dtxsid) %>%
  unique()

cvt %>%
  subset(!(analyte_dtxsid %in% get_cheminfo(info = "DTXSID", model = "gas_pbtk"))) %>%
  dplyr::distinct(analyte_dtxsid, analyte_name_original)

# Test which analyte_dtxsid don't have a clearance or vdist

data.frame(Chemical = unique(cvt$analyte_dtxsid)) %>%
  rowwise() %>%
  mutate(needs_RestCL = ifelse(
    (class(try(calc_total_clearance(dtxsid = Chemical,
                     restrictive.clearance = TRUE,
                     suppress.messages = TRUE))) == "try-error"),
    TRUE, FALSE),
    needs_nonRestCL = ifelse(
      (class(try(calc_total_clearance(dtxsid = Chemical,
                     restrictive.clearance = FALSE,
                     suppress.messages = TRUE))) == "try-error"),
    TRUE, FALSE),
    needs_Vdist = ifelse(
      (class(try(calc_vdist(dtxsid = Chemical,
                     suppress.messages = TRUE))) == "try-error"),
    TRUE, FALSE)) %>%
  ungroup() -> has_all_needed_df

has_all_needed_df %>%
  filter(!(needs_RestCL | needs_nonRestCL | needs_Vdist)) %>%
  pull(Chemical) %>% unique() -> common_chems_dawson

@@

```{r get_clearances}
clearance_df <- data.frame(Chemical = common_chems) %>%
  rowwise() %>%
  mutate(Restrict_Cl = calc_total_clearance(dtxsid = Chemical,
                                            restrictive.clearance = TRUE,
                                            suppress.messages = TRUE),
         nonRestrict_Cl = calc_total_clearance(dtxsid = Chemical,
                                            restrictive.clearance = FALSE,
                                            suppress.messages = TRUE)) %>%
  ungroup()

hep_clearance_df <- data.frame(Chemical = common_chems) %>%
  rowwise() %>%
  mutate(Restrict_Cl = calc_hep_clearance(dtxsid = Chemical,
                                            restrictive.clearance = TRUE,
                                            suppress.messages = TRUE),
         nonRestrict_Cl = calc_hep_clearance(dtxsid = Chemical,
                                            restrictive.clearance = FALSE,
                                            suppress.messages = TRUE)) %>%
  ungroup()

my_tkstats <- eval_tkstats(my_pk) %>%
  dplyr::select(Chemical, Species, CLtot.tkstats, CLtot.nca) %>%
  distinct() %>%
  filter(Chemical %in% common_chems,
         !is.na(CLtot.tkstats), !is.na(CLtot.nca)) %>%
  group_by(Chemical, Species, CLtot.tkstats) %>%
  summarize(CLtot.nca_mean = mean(CLtot.nca),
            CLtot.nca_sd = sd(CLtot.nca))

# Combine these two
combined_clearance_df <- my_tkstats %>% inner_join(clearance_df)

```{r viz}

# Preliminary viz, plot all clearances
combined_clearance_df %>%
  relocate(CLtot.nca_mean, CLtot.nca_sd, .after = tidyselect::last_col()) %>%
  pivot_longer(cols = 3:6,
               names_to = "Clearance_Source",
               values_to = "Value") %>%
  ggplot(aes(x = Value,
             y = Chemical,
             fill = Clearance_Source)) +
  geom_point(shape = 21, alpha = 0.7) +
  scale_x_log10() +
  scale_fill_manual(breaks = c("nonRestrict_Cl", "CLtot.tkstats",
                               "CLtot.nca_mean",
                               "Restrict_Cl"),
                    labels = c("non-restrictive clearance (httk)",
                               "predicted clearance (invivoPKfit)",
                               "mean non-compartmental total clearance (invivoPKfit)",
                               "restrictive clearance (httk)"),
                    values = c("orange2", "grey5", "skyblue3",  "green4")) +
  facet_grid(rows = vars(Species), scales = "free", space = "free") +
  theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        legend.position = "bottom",
        legend.direction = "vertical",
        axis.ticks = element_blank(),
        axis.line = element_blank())

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "tkClearance_comparison_httk_gasPBPK.png"),
       height = 14,
       width = 5,
       device = "png")
```

Finally, what I want to analyze is how good the fits for these chemicals are relative to the rest of the chemicals.

```{r fit_analysis}
wm <- get_winning_model(my_pk)
ingroup_rmse <- rsq(my_pk) %>% inner_join(wm) %>%
  filter(Chemical %in% combined_clearance_df$Chemical) %>%
  mutate(Group = "in httk")

outgroup_rmse <- rsq(my_pk) %>% inner_join(wm) %>%
  filter(!(Chemical %in% combined_clearance_df$Chemical)) %>%
  mutate(Group = "not in httk")

all_group_rmse <- bind_rows(ingroup_rmse, outgroup_rmse)

all_group_rmse %>%
  filter(Species == "rat") %>%
  ggplot(aes(x = Rsq, y = Chemical,
             shape = Group)) +
  geom_point() +
  facet_grid(rows = vars(Group), scales = "free", space = "free") +
  theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        legend.direction = "vertical",
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text.y = element_blank())

ggsave(paste0(Sys.getenv("FIG_DIR"), "Rsq_gaspbtk.png"),
       height = 5,
       width = 5)

```{r httk_solution}
httk::calc_tkstats(dtxsid = unique(combined_clearance_df$Chemical)[[5]])
httk::parameterize_gas_pbtk(restrictive.clearance = TRUE,
                            dtxsid = unique(combined_clearance_df$Chemical)[[5]])

httk::solve_gas_pbtk(dtxsid = unique(combined_clearance_df$Chemical)[[5]],
                     parameters = httk::parameterize_gas_pbtk(
                       restrictive.clearance = TRUE,
                       dtxsid = unique(combined_clearance_df$Chemical)[[5]]
                       ),
                     dose = 10,
                     exp.conc = 0,
                     input.units = "mg",
                     output.units = "mg",
                     days = 3,
                     plots = TRUE)

# For the invivoPKfit functions, we will hold kelim constant
# This constant will be calculated as CLtot/Vdist (or V1, if 2-compartment)
# This constant is derived by httk in the following way:
calc_total_clearance(dtxsid = common_chems[[10]],
                             restrictive.clearance = FALSE) /
  calc_vdist(dtxsid = common_chems[[10]])

```{r running-invivoPKfit}
# This is now updated to use dawson chem table
minimal_pk <- pk(data = cvt %>%
                   filter(analyte_dtxsid %in% common_chems_dawson),
                 mapping = ggplot2::aes(
                   Chemical = analyte_dtxsid,
                   Chemical_Name = analyte_name_original,
                   DTXSID = analyte_dtxsid,
                   CASRN = analyte_casrn,
                   Species = species,
                   Reference = document_id,
                   Media = conc_medium_normalized,
                   Route = administration_route_normalized,
                   Dose = dose_level_normalized,
                   Dose.Units = "mg/kg",
                   Subject_ID = subject_id,
                   Series_ID = series_id,
                   Study_ID = study_id,
                   ConcTime_ID = conc_time_id,
                   N_Subjects = n_subjects_in_series,
                   Weight = weight_kg,
                   Weight.Units = "kg",
                   Time = time_hr,
                   Time.Units = "hours",
                   Value = conc,
                   Value.Units = "mg/L",
                   Value_SD = conc_sd_normalized,
                   LOQ = loq
                 ))

new_pk <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE,
                      suppress.messages = FALSE) +
  settings_optimx(method = c("L-BFGS-B")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  stat_error_model(error_group = vars(Chemical, Species, Reference)) +
  stat_model(model = c("model_1comp", "model_1comp_cl_rest", "model_1comp_cl_nonrest"))
new_pk <- do_data_info(new_pk)

new_pk <- do_prefit(new_pk)

new_pk$prefit$par_DF %>% arrange(Chemical, Species) %>% filter(param_name == "Clint") %>%
  View()

new_pk <- do_fit(new_pk) # For troubleshooting only

new_pk <- do_fit(new_pk, n_cores = 12)

# Multidplyr error because of forgotten browser() call within log_likelihood
new_pk$fit %>% unnest(fit) %>% select(!starts_with("sigma_")) %>% View()

saveRDS(new_pk, "data-raw/someCVT_wHTTK_reform_JOINT.rds")

coef(new_pk) %>% View()
eval_tkstats(new_pk, finite_only = FALSE) %>% View()
get_winning_model(new_pk) %>% View("winning_model")

AIC(new_pk) %>% filter(method %in% "L-BFGS-B") %>% arrange(Chemical, Species) %>%
  View()

pl2comp <- plot(new_pk, use_scale_conc = FALSE,
           n_interp = 10)

pl2comp %>%
  filter(Chemical %in% "DTXSID1043881") %>% pull(final_plot) %>% .[[1]]
```

@@

```{r kelim-questions}

# Original data, filtered for common_chems
# par_DF
my_kelim_prefit <- my_pk$prefit$par_DF %>%
  filter(param_name %in% "kelim",
         Chemical %in% common_chems_dawson) %>%
  distinct(Chemical, Species, param_name, lower_bound, upper_bound, start)

@@

httk_df <- new_pk$prefit$par_DF %>%
  filter(param_name %in% "kelim") %>%
  distinct(Chemical, Species, param_name, start) %>%
  rename(init_restrictive = "start") %>%
  inner_join(
    nonrest_pk$prefit$par_DF %>%
  filter(param_name %in% "kelim") %>%
  distinct(Chemical, Species, param_name, start) %>%
  rename(init_nonrestrictive = "start"))

@@

pl <- plot(new_pk, n_interp = 12, use_scale_conc = FALSE)
pdf(file = paste0(Sys.getenv("FIG_DIR"),
              "AllFits_nonDoseNorm_newPK_reParametizedModels_2024_0412.pdf"),
    height = 6, width = 10)
for (i in 1:nrow(pl)) {
  print(pl$final_plot[[i]])
}
dev.off()

@@

hclust(d = as.dist((1+val_cormat)/2), method = "average") %>% plot()

hclust(d = as.dist((1+val_cormat)/2), method = "average") -> val_hclust

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

@@

```{r library_loading}
devtools::load_all()
library(httk)
library(tidyverse)
my_pk <- readRDS("data-raw/all_cvt_pk.rds")
httk::load_dawson2021()
```

Steps in the analysis:

1. Get chemicals + species combinations in CvTdb that have _httk_ predictions.
2. Safely get parameters, clearance, and solutions for each of these groups.
3. Compare clearances from _httk_ with those from _invivoPKfit_.
4. Compare $f_{up}$ for subset of data with total values available via _httk_.
5. Use new _clearance-aware_ optimization of 1-compartment parameters to
benchmark _httk_ solutions.
6. Estimate the fraction unbound in plasma for chemicals with rat data,
for those same chemicals with human data for Clint and Fup. 
Begin with the assumption that there is restrictive clearance, those chemicals
that do not change when optimized (and produce reasonable fits) are likely to
be _restrictively cleared_, whereas those that increase Fup during maximum
likelihood estimation are _possibly non-restrictively clearaed_. 
7. Other factors such as renal clearance may play a role in those cases.
Validate the clearance estimate by measuring changes in predicted clearance.
Important to identify whether there seems to be a "fraction missing".
8. Wrap the previous steps into attempts to identify true positive
categorizations for Restrictive and Potentially Non-Restrictively metabolized
chemicals (include Lipinski's Rule of 5 too).

@@

```{r dawson}
# try dawson
# can also be run as background job (see data-raw/RC_comparison_bkgd.R)

human_chems_gas <- get_cheminfo(info = "DTXSID",
                                species = "Human",
                                model = "gas_pbtk",
                                median.only = TRUE,
                                physchem.exclude = TRUE,
                                suppress.messages = TRUE)

common_chems_dawson <- cvt %>%
  filter(species %in% c("human", "rat", "mouse"),
         analyte_dtxsid %in% human_chems_gas) %>%
  pull(analyte_dtxsid) %>%
  unique()

rat_chems_gas <- get_cheminfo(info = "DTXSID",
                                species = "Rat",
                                model = "gas_pbtk",
                                median.only = TRUE,
                                physchem.exclude = TRUE,
                                suppress.messages = TRUE)

common_chems_rat <- cvt %>%
  filter(species %in% "rat",
         analyte_dtxsid %in% rat_chems_gas) %>%
  pull(analyte_dtxsid) %>%
  unique()

remove(list = c("human_chems_gas", "rat_chems_gas"))

# How many are in the previous invivoPKfit analysis?
length(common_chems_dawson %in% unique(my_pk$fit %>%
                                         filter(convcode == 0) %>%
                                         pull(Chemical)))

test_parameterization <- function(dtxsid, species = "Human") {
  is(
    tryCatch(
      {
        parameterize_gas_pbtk(dtxsid = dtxsid,
                              species = species,
                              supress.messages = TRUE)  
      }, error = function(e) {
        NULL
      }
    ),
    "list"
  )
}

parameterize_all_gas <- function(dtxsid, species = "Human") {
  # Note that Clint is given in uL/min, which must be converted to L/h
  params <- httk::parameterize_gas_pbtk(dtxsid = dtxsid,
                                        species = species,
                                        default.to.human = FALSE)
  paramdf <- data.frame(Chemical = dtxsid, Species = tolower(species),
                        Qliverf = params$Qliverf,
                        Funbound.plasma = params$Funbound.plasma,
                        Clint = params$Clint*60/1e+6)
  paramdf$fupClint <- paramdf$Funbound.plasma * paramdf$Clint
  paramdf$ExtEff <- paramdf$fupClint/(paramdf$Qliverf + paramdf$fupClint)
  paramdf$Q_Clint <- paramdf$Qliverf/paramdf$Clint
  return(paramdf)
}

if (all(vapply(common_chems_rat, test_parameterization,
               FUN.VALUE = logical(1), species = "Rat"))) {
  rat_params <- do.call(rbind, lapply(common_chems_rat,
                                      parameterize_all_gas,
                                      species = "Rat"))
}

if (all(vapply(common_chems_dawson, test_parameterization,
               FUN.VALUE = logical(1), species = "Human"))) {
  human_params <- do.call(rbind, lapply(common_chems_dawson,
                                        parameterize_all_gas,
                                        species = "Human"))
}
```

```{r get_clearances}
params_dtxsid <- function(dtxsid, species) {
  httk::parameterize_gas_pbtk(dtxsid = dtxsid,
                              species = species,
                              default.to.human = FALSE,
                              suppress.messages = TRUE)
}

make_clearance_df <- function(dtxsid, species) {
  tmp_df <- data.frame(Chemical = dtxsid)
  tmp_params <- lapply(dtxsid, params_dtxsid, species = species)
  tmp_df$restrict_cl <- vapply(tmp_params,
                               \(x) {
                                 calc_total_clearance(
                                   parameters = x,species = species,
                                   suppress.messages = TRUE,
                                   restrictive.clearance = TRUE
                                 )
                               },
                               FUN.VALUE = double(1L)
  )
  tmp_df$nonrestrict_cl <- vapply(tmp_params,
                                  \(x) {
                                    calc_total_clearance(
                                      parameters = x,species = species,
                                      suppress.messages = TRUE,
                                      restrictive.clearance = FALSE
                                    )
                                  },
                                  FUN.VALUE = double(1L)
  )
  return(tmp_df)
}
clearance_df <- make_clearance_df(dtxsid = common_chems_rat, species = "rat")
```

```{r combine_tkstats}
# Start Here if you ran the background job

my_tkstats <- eval_tkstats(my_pk) %>%
  filter(Media %in% "plasma") %>%
  dplyr::select(Chemical, Species, CLtot.tkstats, CLtot.nca) %>%
  distinct() %>%
  filter(
    Species %in% "rat",
    Chemical %in% common_chems_rat,
    !is.na(CLtot.tkstats), !is.na(CLtot.nca)
  ) %>%
  group_by(Chemical, Species, CLtot.tkstats) %>%
  summarize(
    CLtot.nca_mean = mean(CLtot.nca),
    CLtot.nca_sd = sd(CLtot.nca)
  )

# Combine these two
combined_clearance_df <- my_tkstats %>% inner_join(clearance_df)
```

```{r clearance_viz}
combined_clearance_df %>%
  mutate(AD_restrictive = abs(CLtot.tkstats - restrict_cl),
         AD_nonrestrictive = abs(CLtot.tkstats - nonrestrict_cl),
         rough_determination = ifelse(
           min(AD_restrictive, AD_nonrestrictive) == AD_restrictive &
             CLtot.tkstats < nonrestrict_cl,
           "Restrictive", "Possibly non-Restrictive"
         )) %>% 
  dplyr::distinct(Chemical, rough_determination) %>%
  View()

# Preliminary viz, plot all clearances
ccd <- combined_clearance_df %>%
  relocate(CLtot.nca_mean, CLtot.nca_sd, .after = tidyselect::last_col()) %>%
  pivot_longer(
    cols = 3:6,
    names_to = "Clearance_Source",
    values_to = "Value"
  ) %>%
  mutate(
    Chemical = factor(Chemical),
    Clearance_Source = factor(
      Clearance_Source,
      levels = c(
        "CLtot.tkstats", "CLtot.nca_mean",
        "nonrestrict_cl", "restrict_cl"
      )
    )
  ) %>%
  transform(
    Chemical = fct_reorder(Chemical, Value, max)
  ) %>% 
  as_tibble()

ccd %>%
  ggplot(aes(
    x = Value,
    y = Chemical,
    fill = Clearance_Source
  )) +
  geom_point(shape = 21, alpha = 0.7) +
  scale_x_log10() +
  scale_fill_manual(
    breaks = c(
      "nonrestrict_cl", "CLtot.tkstats",
      "CLtot.nca_mean",
      "restrict_cl"
    ),
    labels = c(
      "non-restrictive clearance (httk)",
      "predicted clearance (invivoPKfit)",
      "mean non-compartmental total clearance (invivoPKfit)",
      "restrictive clearance (httk)"
    ),
    values = c("orange2", "grey5", "skyblue3", "green4")
  ) +
  facet_grid(rows = vars(Species), scales = "free", space = "free") +
  labs(
    x = "Total Clearance",
    y = "Chemical"
  ) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1.5),
    panel.background = element_blank(),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", angle = 0),
    # axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "right",
    legend.direction = "vertical",
    axis.ticks = element_blank(),
    axis.line = element_blank()
  )

clear_chem_levels <- levels(ccd$Chemical)

ggsave(
  filename = paste0(
    Sys.getenv("FIG_DIR"),
    "rat_tkClearance_comparison_httk_PBPK.png"
  ),
  height = 8,
  width = 8,
  device = "png"
)
```

I would also like to know, how does the $f_{up}$ of this set of chemicals
compare to that of all chemicals in `httk` (Dawson 2021).

```{r fup_comparison}
fup_df <- data.frame(
  Chemical = get_cheminfo("DTXSID", species = "Rat"),
  fup = as.numeric(vapply(get_cheminfo("Funbound.plasma", species = "Rat"),
                          gsub, pattern = ",.+", replacement = "",
                          FUN.VALUE = character(1)
  ))
)

fup_df %>%
  ggplot(aes(x = fup)) +
  geom_histogram(
    # aes(y = after_stat(density)),
    binwidth = 0.03,
    color = "grey5", fill = "grey5"
  ) +
  geom_histogram(
    data = fup_df %>%
      filter(Chemical %in% common_chems_rat),
    # aes(y = after_stat(density)),
    fill = NA,
    color = "darkgreen", binwidth = 0.03
  ) +
  expand_limits(expand = c(0,0)) +
  # geom_vline(xintercept = 0.05, linetype = "dashed") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  labs(
    title = "Distribution of Funbound.plasma values",
    subtitle = "(chemicals fit by invivoPKfit in dark green)"
  ) +
  coord_cartesian(clip = "off") +
  theme_bw()

ggsave(paste0(Sys.getenv("FIG_DIR"), "fup_density_plots.png"),
       height = 5,
       width = 6,
       device = ragg::agg_png,
       units = "in"
)

test_fup <- fup_df %>%
  # filter(Chemical %in% common_chems_dawson) %>%
  pull(fup)
for (i in seq(0, 0.5, by = 1e-9)) {
  if (sum(test_fup > i) == sum(test_fup <= i)) {
    print(i)
    break
  }
}
```

Note that this is only for rats. This is supposed to be the best case,
but of course _httk_ is "made" for predicting **human exposures**. Also,
I will try to order some of this data the same way as the clearance figures.

This will ultimately help understand whether Total Clearance
differences and prediction differences relate. Of course it would
be possible to do a formal test for correlation but that overdoes what we need
to build this relationship for with the rather limited _rat_ data we have
in both _httk_ and _invivoPKfit_.

```{r rat_invivo_comparison}

solve_gas_pbtk(
  dtxsid = common_chems_dawson[10],
  dose = 1, exp.conc = 0,
  iv.dose = FALSE,
  input.units = "mg/kg",
  output.units = "mg/L",
  suppress.messages = TRUE
) %>%
  ggplot(aes(x = time)) +
  geom_line(aes(y = Cplasma)) +
  geom_line(aes(y = Crest), color = "blue4") +
  geom_line(aes(y = Cgut), color = "goldenrod") +
  geom_line(aes(y = Cliver), color = "darkgreen") +
  geom_line(aes(y = Clung), color = "hotpink") +
  geom_line(aes(y = Ckidney), color = "purple4")

# Functions wrapping httk single dose evaluation
httk_bolus_gas_pbtk <- function(dtxsid, species = "Human",
                                dose, times, iv.dose, LOQ) {
  tmp_solution <- solve_gas_pbtk(
    dtxsid = unique(dtxsid),
    dose = unique(dose),
    exp.conc = 0,
    iv.dose = unique(iv.dose),
    times = unique(c(0, times/24)),
    input.units = "mg/kg",
    output.units = "mg/L",
    species = unique(stringr::str_to_title(species)),
    default.to.human = TRUE,
    suppress.messages = TRUE
  )
  # Remove the 0.0001 intermediate time value that is returned
  retdf <- tibble(
    Time = round(tmp_solution[, "time"]*24, digits = 3),
    httk_Preds = tmp_solution[, "Cplasma"]
  )
  retdf[which(retdf$Time %in% unique(round(times, 3))), ]
}
```

We need to put the _in vivo_ data in an amenable format.
For this I will limit the species to "rat", the chemicals to
`common_chems_rat`, and the medium to "plasma".

```{r rat_data}
my_data <- get_data(my_pk)

rat_df <- my_data %>%
  distinct(Chemical, Species, Dose, Route, Media, Time, Conc, exclude, Detect,
           pLOQ) %>%
  filter(Species %in% "rat", Media %in% "plasma",
         Chemical %in% common_chems_rat, !exclude, Detect) %>%
  mutate(iv_dose = (Route == "iv"),
         Time = round(Time, 3)) %>%
  group_by(Chemical, Species, Dose, Route, Media) %>%
  filter(n() > 4) %>%
  nest(.key = "CvT") %>%
  rowwise() %>%
  mutate(
    httk_solved = list(
      tryCatch(
        expr = do.call(httk_bolus_gas_pbtk,
                       list(
                         dtxsid = Chemical,
                         species = Species,
                         dose = Dose,
                         times = CvT$Time,
                         iv.dose = CvT$iv_dose,
                         LOQ = CvT$pLOQ
                       )
        ) %>%
          inner_join(tibble(Time = CvT$Time, Conc = CvT$Conc, pLOQ = CvT$pLOQ),
                     by = join_by(Time)),
        error = function(err) {
          message(paste0("Unable to get solution for ", Chemical))
          print(err)
        } 
      )
    )
  ) %>% filter(Chemical %in% clear_chem_levels)

length(unique(rat_df$Chemical))

rat_df$Chemical <- factor(rat_df$Chemical, levels = clear_chem_levels)

rat_df <- rat_df %>%
  dplyr::select(Chemical, Species, Dose, Route, Media, httk_solved) %>%
  unnest(cols = httk_solved)

rat_out <- rat_df %>%
  mutate(httk_Preds = ifelse(httk_Preds < pLOQ, pLOQ, httk_Preds)) %>% 
  mutate(SLE = (log2(Conc) - log2(httk_Preds))^2,
         AE = abs(Conc - httk_Preds)) %>%
  summarize(
    RMSLE = sqrt(mean(SLE, na.rm = TRUE)),
    MAE = mean(AE, na.rm = TRUE)
   ) %>% pivot_longer(cols = c(RMSLE, MAE),
                      names_to = "metric",
                      values_to = "value")

rat_out %>%
  na.omit() %>%
  ggplot() +
  geom_tile(data = rat_out %>% na.omit() %>%
              filter(metric %in% "RMSLE"),
            aes(x = metric, y = Chemical, fill = value)) +
  scale_fill_gradient(low = "grey5", high = "grey90",
                      name = "RMSLE") +
  ggnewscale::new_scale_fill() +
  geom_tile(data = rat_out %>% na.omit() %>%
              filter(metric %in% "MAE"),
            aes(x = metric, y = Chemical, fill = value)) +
  scale_fill_gradient(low = "darkgreen", high = "seagreen1",
                      limits = c(0, 80),
                      labels = c("0", "20",
                                 "40", "60",
                                 "80+"),
                      oob = scales::squish,
                      name = "MAE") +
  theme_minimal() +
  theme(axis.text.x = element_text(hjust = 1, angle = 50),
        legend.title = element_text(hjust = 0.5)) +
  coord_fixed(ratio = 0.75)

ggsave(
  filename = paste0(
    Sys.getenv("FIG_DIR"),
    "rat_httk_predictionRMSLE.png"
  ),
  height = 8.1,
  width = 4,
  device = "png"
)

rat_out2 <- rat_df %>%
  group_by(Chemical, Species, Route, Dose) %>% nest() %>%
  summarize(plot = map(
    data,
    \(d){
      ggplot(d, aes(x = Time)) +
        geom_point(aes(y = Conc)) +
        geom_line(aes(y = httk_Preds),
                    color = "forestgreen") +
        geom_point(aes(y = httk_Preds),
                   color = "forestgreen",
                   shape = 1) +
        labs(title = Chemical,
             subtitle = paste(Dose, "mg/kg given via",
                               Route, "bolus"),
             y = "Plasma Concentration (mg/L)",
             x = "Time (hrs)") +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5),
              plot.subtitle = element_text(hjust = 0.5))
    }
  ))
```

```{r httk_solution}

rat_params %>%
  ggplot(aes(y = fupClint)) + 
  geom_point(aes(x = Clint), color = "darkred", shape = 1) +
  geom_abline(slope = 1, intercept = 0, color = "darkblue",
              linetype = "dashed")

rat_params %>%
  mutate(fup_z = scale(Funbound.plasma),
         clint_z = scale(Clint)) %>%
  dplyr::select(Chemical, fup_z, clint_z) %>%
  pivot_longer(cols = c(fup_z, clint_z),
               names_to = "pars", values_to = "vals") %>%
  ggplot(aes(x = pars, y = vals, group = Chemical)) +
  geom_line()

httk::calc_tkstats(dtxsid = unique(rat_out$Chemical)[[5]])
httk::parameterize_gas_pbtk(
  restrictive.clearance = TRUE,
  dtxsid = unique(rat_out$Chemical)[[5]]
)

httk::solve_gas_pbtk(
  dtxsid = unique(combined_clearance_df$Chemical)[[5]],
  parameters = httk::parameterize_gas_pbtk(
    restrictive.clearance = TRUE,
    dtxsid = unique(combined_clearance_df$Chemical)[[5]]
  ),
  dose = 10,
  exp.conc = 0,
  input.units = "mg",
  output.units = "mg",
  days = 3,
  plots = TRUE
)

# For the invivoPKfit functions, we will hold kelim constant
# This constant will be calculated as CLtot/Vdist (or V1, if 2-compartment)
# This constant is derived by httk in the following way:
calc_total_clearance(
  dtxsid = common_chems[[10]],
  restrictive.clearance = FALSE
) /
  calc_vdist(dtxsid = common_chems[[10]])
```

```{r running-invivoPKfit}
# This is now updated to use dawson chem table
minimal_pk <- pk(data = cvt %>%
                   filter(analyte_dtxsid %in% common_chems_rat,
                          species %in% "rat"))

new_pk <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(
    keep_data_original = TRUE,
    suppress.messages = FALSE
  ) +
  settings_optimx(method = c("L-BFGS-B")) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  stat_error_model(error_group = vars(Chemical, Species)) +
  stat_model(model = c(
    "model_1comp", "model_1comp_cl_rest", 
    "model_1comp_cl_nonrest", "model_1comp_fup",
    "model_flat")) +
  settings_optimx(method = c("bobyqa"))

new_pk <- do_data_info(new_pk)

new_pk <- do_prefit(new_pk)
gc()

unique(new_pk$prefit$par_DF$param_name)

new_pk$prefit$par_DF %>%
  arrange(Chemical, Species) %>%
  filter(param_name == "Fup") %>%
  View()

new_pk <- do_fit(new_pk, n_cores = 14)
gc()

new_pk$fit %>%
  View()

saveRDS(new_pk, "data-raw/ratCVT_wHTTK_reformulated.rds")
```

```{r begin_analysis}
new_pk <- readRDS("data-raw/ratCVT_wHTTK_reformulated.rds")

coef(new_pk) %>% View()
eval_tkstats(new_pk, finite_only = FALSE) %>% View()
get_winning_model(new_pk) %>%
  count(model)

AIC(new_pk) %>%
  arrange(Chemical, Species) %>%
  group_by(model, method) %>% 
  count(is.finite(AIC))

pl2comp <- plot(new_pk,
                use_scale_conc = FALSE,
                n_interp = 12
)

pl2comp %>%
  filter(Chemical %in% "DTXSID4020533") %>%
  pull(final_plot)

rel_ll <- AIC(new_pk) %>% 
  filter(is.finite(AIC)) %>% 
  dplyr::select(-c(log_likelihood, npar)) %>% 
  pivot_wider(names_from = model,
              values_from = AIC) %>% 
  rowwise() %>% 
  mutate(minAIC = min(c_across(starts_with("model_")), na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(across(starts_with("model_"), \(x) {
    signif(exp((minAIC - x)/2) * 100, 4)
  }))

# Test whether the flat model has lowest relative likelihood
rel_ll %>%
  dplyr::select(-c(Species, method))
# We can inspect this visually
rel_ll %>% arrange(desc(model_flat)) %>% head()
rel_ll %>%
  dplyr::select(-c(Species, method)) %>% 
  rowwise() %>% 
  mutate(across(contains("_1comp"),
                \(x) {
                  ifelse(
                    x < model_flat | x < 1,
                    NA_real_,
                    x
                  )
                })) %>% 
  filter(!if_all(contains("_1comp_"), is.na)) %>%
  clipr::write_clip()

rel_ll %>%
  rowwise() %>% 
  mutate(across(contains("_1comp"),
                \(x) {
                  ifelse(
                    x < model_flat | x < 1,
                    NA_real_,
                    x
                  )
                })) %>% 
  drop_na(model_1comp_fup)

new_pk$fit %>%
  filter(param_name %in% "Fup",
         model %in% "model_1comp_fup") %>%
  ggplot(aes(x = estimate, y = start)) +
  geom_point(size = 2) +
  theme_bw() +
  coord_equal(xlim = c(0,1), ylim = c(0,1),
              expand = TRUE)

new_pk$fit %>%
  filter(param_name %in% "Fup",
         model %in% "model_1comp_fup") %>%
  distinct(Chemical, Species, method, model,
           param_name, estimate, start) %>%
  mutate(direction = ifelse(abs(estimate - start) < 0.1,
                            "similar (less than 10% difference)",
                            ifelse(estimate > start,
                                   "increase",
                                   "decrease"))) %>%
  pivot_longer(cols = c(estimate, start),
               names_to = "parameterization",
               values_to = "value") %>%
  mutate(parameterization = factor(parameterization,
                                   levels = c("start", "estimate"))) %>%
  ggplot(aes(x = parameterization, y = value)) +
  geom_point() +
  geom_line(aes(group = Chemical)) +
  facet_grid(cols = vars(direction)) +
  labs(title = "Optimization of fraction unbound in plasma") +
  theme_bw()

ggsave(filename = paste0(
  Sys.getenv("FIG_DIR"),
  format(today(), '%d%b%Y'),
  "fup_optimization_generalFits"
))

@@

@@

```{r kelim-questions}
# Original data, filtered for common_chems
# par_DF
my_kelim_prefit <- my_pk$prefit$par_DF %>%
  filter(
    param_name %in% "kelim",
    Chemical %in% common_chems_dawson
  ) %>%
  distinct(Chemical, Species, param_name, lower_bound, upper_bound, start)

@@

httk_df <- new_pk$prefit$par_DF %>%
  filter(param_name %in% "kelim") %>%
  distinct(Chemical, Species, param_name, start) %>%
  rename(init_restrictive = "start") %>%
  inner_join(
    nonrest_pk$prefit$par_DF %>%
      filter(param_name %in% "kelim") %>%
      distinct(Chemical, Species, param_name, start) %>%
      rename(init_nonrestrictive = "start")
  )

@@

pl <- plot(new_pk, n_interp = 12, use_scale_conc = FALSE)
pdf(
  file = paste0(
    Sys.getenv("FIG_DIR"),
    "AllFits_nonDoseNorm_newPK_reParametizedModels_2024_0412.pdf"
  ),
  height = 6, width = 10
)
for (i in seq_len(nrow(pl))) {
  print(pl$final_plot[[i]])
}
dev.off()
```

@@

hclust(d = as.dist((1 + val_cormat) / 2), method = "average") %>% plot()

hclust(d = as.dist((1 + val_cormat) / 2), method = "average") -> val_hclust
